### 2018-04-25 Docs/Thesis/docs/chapter3.tex
update
thesis update

In this chapter, we will present our tool which aims to fill the gaps addressed in the chapter ~\ref{ch:introduction}. We call it \textit{SIMITAR}, an acronym for \textit{SnIffing, ModellIng and TrAffic geneRation}.  This acronym summarizes its operation. SIMITAR is a traffic generator able to learn features of a real traffic automatically, and reproduce a synthetic traffic similar to the original. It records a model for the traffic in an XML file we call \textit{compact trace descriptor} (CTD file). As input-data SIMITAR can use \textit{pcap} files or real-time captures.

@@

Our primary goal is to offer simple configuration, with \textbf{realism}, at a reasonable speed. The intermediate layer of the figure ~\ref{fig:layer-diagram} summarize, the goal of the project in an illustrative way. Instead of storing a huge file \textit{pcap} with a size of many \textit{gigas}, we create a light-weight set of models and parameters that describe this same trace. We are also introducing the concept of \textbf{programmability}. Since the \textit{compact trace descriptor},  is an XML file, the user may create his custom traffic in a platform agnostic way creating his own CTD file. So the will be able to create multi-flow Ethernet workload not having to create many scripts for each tool or reading its documentation. A single description of the traffic will fit any supported packet generator. Using a component methodology, we decouple the packet generation, from the data collection and parameterization process. We developt it using the design pattern factory to make the extension easy for any packet generator engine.

@@

It has a data structure we developed called \textit{OrderedSet}. A set is a list of elements with no repetition but does not keep track of the insertion order. Our OrderedSet does. Also, it makes use of a  64 bits hash function of the family FNV\footnote{The collision probability of a good 64 bits hash function in a table with 10000 items is about of $2.71e-12$.}. The listed header fields are inputs for a hash function, and its value is set on the ordered set which returns its order (index on the \textit{OrderedSet}). The index value is chosen as a packet flowID.

@@

Some unique per-flow features are directly measured from the data. They are:

@@

Each one of these parameters is unique per flow. Other features like PSD (packet size distribution) e IPT (Inter-packet time), have a more complex behavior.  To represent these characteristics, we will use sets of stochastic-based models.

@@

To represent inter-packet times, we adopt a simplified version of the Harpoon's traffic model. A deep explanation of the original model can be found at \cite{harpoon-paper} and \cite{harpoon-validation}. Here, we will explain our simplified version, which is illustrated at figure~\ref{fig:modified-harpoon-model}. Harpoon uses a definition of each level, based on the measurement of SYN and ACK TCP flags. It uses TCP flags (SYN) to classify packets in different levels, named their file, session, and user level. We choose to estimate these values, based on inter-packet times only. The distinction is made based on the time delay between packets.

In our algorithm simplified version, we define three different layers of data transference to model and control: file, session, and flow. For SIMITAR, a file is just a sequence of consecutive packets transmitted continuously, without large interruption of traffic. It can be, for example, packets sent downloading a file, packets from a  UDP  connection or a single ICMP echo packet. The session-layer refers to a sequence of multiple files transmitted between a source and a destination, belonging to the same flow.  The flow level refers to the conjunct of flows, as classified by the Sniffer.  Now, we explain SIMITAR operation on each layer.

@@

In the \textbf{file layer}, we model the inter-packet times at the file level. To estimate inter-packet times within files, we select all inter-packet times smaller than \texttt{session\_cut\_time}\footnote{In the code it is called\texttt{DataProcessor::m\_session\_cut\_time} }. All files within the same flow are considered to follow the same model. We delegate the control of the inter-packet times to the underlying packet generator engine. We ordered them, from the best to the worst. Currently, we are using eight different stochastic functions parameterizations. They are Weibull(linear regression), Normal(mean/standard deviation calculation), Exponential(mean and linear regression estimation), Pareto(linear regression and maximum likelihood), Cauchy(linear regression) and Constant(mean calculation). From those, Weibull, Pareto, and Cauchy are heavy-tailed functions, and therefore self-similar processes. But if the flow has less than 30 packets, just the constant model is evaluated. It is because numerical methods gave poor results if the data sample used is small. We sort these models according to the Akaike Information Criterion (AIC) as default\cite{sourcesonoff-paper}\cite{bic-aic-comparision}. We will enter into deeper details on this methodology on the chapter ~\ref{ch:modeling-evaluation}. The methodology of selection is presented in the figure~\ref{fig:model-parameterization}, and all constants and modes of operation can be changed by command line options.

@@

Our approach for the packet size is much simpler. Since the majority of packet size distribution found on real measurements are bi-modal \cite{packet-distribution-model}\cite{sourcesonoff-paper}\cite{udp-flows-model}, we first sort all packet sizes of flow in two modes. We define a packet size mode cut value of 750 bytes, same value adopted by \cite{udp-flows-model}.

We know how much packets each mode has, and then we fit a model for it. We use three stochastic models: constant, exponential and normal. Since self-similarity does not make sense for packet-sizes, we prefer to use just the simpler models. When there is no packet for a model, we set a flag NO\_MODEL, and when there is just a single packet we just use the constant model. Then calculate the BIC and AIC for each, but we decide to set the constant model as the first.

@@

\subsubsection{Compact Trace Descriptor}
%<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<

@@

The Flow Generator handles the data on the \textit{Compact Trace Descriptor} file, and use to retrieve parameters for traffic generation. It crafts and controls each flow in a separated thread. We already implemented this component using Iperf and Libtins(C++ API) as packet generators. It must follow the class hierarcy as presented in the figure ~\ref{fig:network-trace-flow-class-diagram}.  This component was designed using the design pattern Factory, to simplify its expansion and support\footnote{If the user wants to introduce support for a new packet generator engine, he just has to expand the class \texttt{DummyFlow}, creating a new derived class (figure ~\ref{fig:network-trace-flow-class-diagram} ). On our current implementation we already have implemented \texttt{IperfFlow} and \texttt{TinsFlow}, and \texttt{DitgFlow} is under validation process. This new class has the requirement of not have any new attribute. The support for this new class must be given on the factory class \texttt{NetworkFlowFactory}. For closed loop packet-crafters (that means, tools which must establish a connection between the source and the destination), two methods must be implemented: \texttt{flowGenerate()} and \texttt{server()}. \texttt{flowGenerate()} is responsible for sending a \textit{file}, as defined on de figure~\ref{fig:modified-harpoon-model}, and \texttt{server()} to receive $n$ files. Open loop packet-crafters, such as we implemented using Libtins (just sent the packets, do not establishes a connection), the server do not need to be implemented.}.

This component itself is a multi-layer workload generator according to the typing introduced on chapter ~\ref{ch:literature-review} \footnote{Since it works both at packet and flow level, but do not work at application level.}. At the flow-level, SIMITAR controls each flow using the algorithm at figure~\ref{fig:alg-simplified-harpoon}. This algorithm handles our model defined at the figure ~\ref{fig:modified-harpoon-model}. This procedure is independent of the underlying packet crafting tool used. It starts a \textit{thread} for each flow in the \textit{Compact Trace Descriptor}, then the thread sleeps for \texttt{start\_delay} seconds. This is the arrival time of the first flow packet. Passed this time, it then calls the underlying packet generator tool(defined as command line argument), and pass to it the flowID, file ON time, number of packets and number of bytes to be sent (file size), and network interface. Then it sleeps the next session OFF time. When the list of ON/OFF times from the flow is over, the thread ends.

At the packet level, SIMITAR configures the packet-generator tool to send a \textit{file}, as defined in the figure ~\ref{fig:modified-harpoon-model}. This method must use the available parameters, attributes and \textit{getters} to configure and generate a thread-safe traffic using the method \texttt{flowGenerate()}. The \textit{file} configuration must follow the in the figure ~\ref{fig:alg-traffic-engine-config}.  Down below we present a simple code of how the D-ITG API can be used to generate packet-level traffic. A more complex configuration is possible, but it serves to illustrate the procedure. We call this concept \textit{Flow Generation Programming}.  Its API documentation is available at on D-ITG website \footnote{\href{http://www.grid.unina.it/software/ITG/manual/index.html\#SECTION00047000000000000000}{http://www.grid.unina.it/software/ITG/manual/index.html\#SECTION00047000000000000000}}.

@@

A network packet generator is a tool or lybrary that should provide its API or script interface\footnote{In the case of a tool with a available CLI, to use it as packet generator we must use functions such as \texttt{fork()} or \texttt{popen()}. For Iperf, we used \texttt{popen()}.} for the \textit{FlowGenerator} component. With this engine, the user must be eble to send packets and controll atributes such as sending time, bandwidth, number of packets, protocols, and so on. That means, any available parameter form the compact trace descriptor.

@@

# @ SIMITAR/trace-analyzer/, execute this command to create two Compact Trace Descriptors, called intrig.ms.xml and intrig.sec.xml. The first is parameterized using milisecons, and de second, uses seconds as time unity.
./trace-analyzer --trace intrig

@@

An example of use case is presented in the figure ~\ref{fig:use-case}. A Device under test can be stressed using a combination of traffic generated by many clients and server pairs. In the case of an open-loop packet generator tool (such as in the Libtins implementation), the servers pairs are not necessary. Using passive network measures, such as Pathload and pathChirp \cite{swing-paper}, it is possible to measure statistics form the device under tests.

>>>>

In this chapter, we will present our tool which aims to fill the gaps addressed in the chapter ~\ref{ch:introduction}. We call it \textit{SIMITAR}, an acronym for \textit{SnIffing, ModellIng and TrAffic geneRation}.  This acronym summarizes its operation. SIMITAR is a traffic generator able to learn features of real traffic automatically, and reproduce synthetic traffic similar to the original. It records a model for the traffic in an XML file we call \textit{compact trace descriptor} (CTD file). As input-data SIMITAR can use \textit{pcap} files or real-time captures.

@@

Our primary goal is to offer simple configuration, with \textbf{realism}, at a reasonable speed. The intermediate layer of the figure ~\ref{fig:layer-diagram} summarize, the purpose of the project in an illustrative way. Instead of storing a huge file \textit{pcap} with a size of many \textit{gigas}, we create a light-weight set of models and parameters that describe this same trace. We are also introducing the concept of \textbf{programmability}. Since the \textit{compact trace descriptor},  is an XML file, the user may create his custom traffic in a platform agnostic way creating his own CTD file. So the user will be able to create multi-flow Ethernet traffic not having to build many scripts for each tool or reading its documentation. A single description of the traffic will fit any supported packet generator. Using a component methodology, we decouple the packet generation, from the data collection and parameterization process. We developt it using the design pattern factory to make the extension easy for any packet generator engine.

@@

It has a data structure we developed called \textit{OrderedSet}. A set is a list of elements with no repetition but does not keep track of the insertion order. Our OrderedSet does. Also, it makes use of a  64 bits hash function of the family FNV\footnote{The collision probability of a good 64 bits hash function in a table with 10000 items is about of $2.71e-12$.}. The listed header fields are inputs for a hash function. The hash value is then added to the ordered set which returns its order (index on the \textit{OrderedSet}). The index value is chosen as a packet flowID.

@@

We measure some flow-features directly from data. They are:

@@

Each one of these parameters is unique per flow. Other features like packet-size distribution and inter-packet times follow probability distributions.  To represent these characteristics, we will use sets of stochastic-based models.

@@

To represent inter-packet times, we adopt a modified version of the Harpoon's traffic model. A deep explanation of the original model can be found at \cite{harpoon-paper} and \cite{harpoon-validation}. Harpoon uses a definition of each level, based on the measurement of SYN and ACK TCP flags. It uses TCP flags (SYN) to classify packets in different levels, named their file, session, and user level. We choose to estimate these values, based on inter-packet times only. The distinction is made based on the time delay between packets. For our purposes, the advantage of this strategy is that the evaluation of the packet-train is not attached to a specific protocol. We illustrate this behavior in the figure~\ref{fig:modified-harpoon-model}.

In our algorithm, we define three different layers of data transference to model and control: \textit{file}, \textit{session}, and \textit{flow}. For SIMITAR, a \textit{file} is a sequence of consecutive packets transmitted continuously, without long interruption of traffic. A file can be, for example, packets from download, a UDP   connection or a single ICMP echo packet. The session-layer refers to a sequence of multiple files transmitted between a source and a destination, belonging to the same flow.  The flow level refers to the conjunct of flows, as classified by the Sniffer.  Now, we explain SIMITAR operation on each layer.

@@

In the \textbf{file layer}, we model the inter-packet times at the file level. To estimate inter-packet times within files, we select all inter-packet times smaller than \texttt{session\_cut\_time}\footnote{In the code it is called\texttt{DataProcessor::m\_session\_cut\_time} }. All files within the same flow are considered to follow the same model. We delegate the control of the inter-packet times to the underlying packet generator engine. We ordered them, from the best to the worst. Currently, we are using eight different stochastic functions parameterizations. They are Weibull(linear regression), Normal(mean/standard deviation calculation), Exponential(mean and linear regression estimation), Pareto(linear regression and maximum likelihood), Cauchy(linear regression) and Constant(mean calculation). From those, Weibull, Pareto, and Cauchy are heavy-tailed functions, and therefore self-similar processes. But if the flow has less than 30 packets, just the constant model is evaluated. It is because numerical methods gave poor results if the data sample used is small. We sort these models according to the Akaike Information Criterion (AIC) as default\cite{sourcesonoff-paper}\cite{bic-aic-comparision}. We will enter into deeper details on this methodology on the chapter ~\ref{ch:modeling-evaluation}. We present the methodology of selection in the figure~\ref{fig:model-parameterization}. All these constants and modes of operation are modifiable via command-line options.

@@

Our approach for the packet size is much simpler. Since the majority of packet size distribution found on real measurements are bi-modal \cite{packet-distribution-model}\cite{sourcesonoff-paper}\cite{udp-flows-model}, we first sort all packet sizes of flow in two modes. We define a packet-size cut value of 750 bytes, same value adopted by \cite{udp-flows-model}.

We know how much packets each mode has, and then we fit a model for it. We use three stochastic models: constant, exponential and normal. Since self-similarity does not make sense for packet-sizes, we prefer to use just the simpler models. When there is no packet for a model, we set a flag NO\_MODEL, and when there is only a single packet, we use the constant model. Then calculate the BIC and AIC for each, but we decide to set the constant model as the first.

@@

\subsubsection{Compact Trace Descriptor}

@@

The Flow Generator handles the data on the \textit{Compact Trace Descriptor} file, and use to retrieve parameters for traffic generation. It crafts and controls each flow in a separated thread. We already implemented this component using Iperf and Libtins(C++ API) as packet generators. It must follow the class hierarchy as presented in the figure ~\ref{fig:network-trace-flow-class-diagram}.  This component was designed using the design pattern Factory, to simplify its expansion and support\footnote{If the user wants to introduce support for a new packet generator engine, he has to expand the class \texttt{DummyFlow}, creating a new derived class (figure ~\ref{fig:network-trace-flow-class-diagram} ). On our current implementation we already have implemented \texttt{IperfFlow} and \texttt{TinsFlow}, and \texttt{DitgFlow} is under validation process. This new class has the requirement of not have any new attribute. The support for this new class must be implemented on the factory class \texttt{NetworkFlowFactory}. For closed loop packet-crafters (that means, tools which must establish a connection between the source and the destination), two methods must be implemented: \texttt{flowGenerate()} and \texttt{server()}. \texttt{flowGenerate()} is responsible for sending a \textit{file}, as defined on de figure~\ref{fig:modified-harpoon-model}, and \texttt{server()} to receive $n$ files. For packet-crafter that support open-loop traffic\footnote{just sent the packets, do not establishes a connection}, such as the one we implemented using Libtins, the server side not needs to be implemented.}.

This component itself is a multi-layer workload generator according to the typing introduced on chapter ~\ref{ch:literature-review} \footnote{Since it works both at packet and flow level, but do not work at application level.}. At the flow-level, SIMITAR controls each flow using the algorithm at figure~\ref{fig:alg-simplified-harpoon}. This algorithm handles our model defined in the figure ~\ref{fig:modified-harpoon-model}. This procedure is independent of the underlying packet crafting tool used. It starts a \textit{thread} for each flow in the \textit{Compact Trace Descriptor}, then the thread sleeps for \texttt{start\_delay} seconds. The \texttt{start\_delay} is the arrival time of the first flow packet. Passed this time, it then calls the underlying packet generator tool(defined as command line argument), and pass to it the flowID, file ON time, number of packets and number of bytes to be sent (file size), and network interface. Then it sleeps the next session OFF time. When the list of ON/OFF times from the flow is over, the thread ends.

At the packet level, SIMITAR configures the packet-generator tool to send a \textit{file}, as defined in the figure ~\ref{fig:modified-harpoon-model}. This method must use the available parameters, attributes and \textit{getters} to configure and generate thread-safe traffic using the method \texttt{flowGenerate()}. The \textit{file} configuration must follow the in the figure ~\ref{fig:alg-traffic-engine-config}.  Down below we present a simple code of how the D-ITG API can be used to generate packet-level traffic. A more complex configuration is possible, but it serves to illustrate the procedure. We call this concept \textit{Flow Generation Programming}.  Its API documentation is available at on D-ITG website \footnote{\href{http://www.grid.unina.it/software/ITG/manual/index.html\#SECTION00047000000000000000}{http://www.grid.unina.it/software/ITG/manual/index.html\#SECTION00047000000000000000}}.

@@

A network packet generator is a tool or library that should provide its API or script interface\footnote{In the case of a tool with an available CLI, to use it as packet generator we must use functions such as \texttt{fork()} or \texttt{popen()}. For Iperf, we used \texttt{popen()}.} for the \textit{FlowGenerator} component. With this engine, the user must be able to send packets and control attributes such as sending time, bandwidth, number of packets, protocols, and so on. That means, any available parameter form the compact trace descriptor.

@@

# @ SIMITAR/trace-analyzer/, execute this command to create two Compact Trace Descriptors, called intrig.ms.xml and intrig.sec.xml. The first is parameterized using milliseconds, and de second uses seconds as time unity.
./trace-analyzer --trace intrig

@@

An example of use case is presented in the figure ~\ref{fig:use-case}. A Device under test can be stressed using a combination of traffic generated by many clients and server pairs. In the case of an open-loop packet generator tool (such as in the Libtins implementation), the servers and clients pairs are not necessary. Using passive network measures, such as Pathload and pathChirp \cite{swing-paper}, it is possible to measure statistics form the device under tests.
